---
title: "John_Day_River"
author: "Markus Min"
date: "1/5/2022"
output: html_document
---

In this Rmd we will be recreating Shelby Richins' results for John Day River Steelhead.

NOTE: PIT tag detectors were only installed in the two adult fish ladders at John Day Dam in 2017. So from 2005-2015 we last see them at the Dalles Dam, and then in the John Day River


# PTAGIS Query

## Query 1 - adult returns to BON
- Interrogation Detail

Query Builder2:
- 1 Attributes: Standard
- 2 Metrics: Obs Count, Obs Time Max, Obs Time Min
- 9 Mark Site Subbasin == Upper John Day, North Fork John Day, Middle Fork John Day, Lower John Day
  - Searched "John Day" and selected all
- 16 Obs Time: Value between 6/1/2005 12:00:00 AM and 5/31/2015 11:59:59
- 17 Obs Site: BO1 - Bonneville Bradford Is. Ladder, BO2 - Bonneville Cascades Is. Ladder, BO3 - Bonneville WA Shore Ladder/AFF, BO4 - Bonneville WA Ladder Slots
- 24 Species: Steelhead
- 29 Mark Length: Less than or equal to 350


### Examine data

#### Load libraries
```{r load_libraries}
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
```

#### Import data
```{r load_data}
JDR_BON_adult_returns <- clean_names(read.csv(here("PTAGIS_queries", "intermediate_files",  "2022-01-06-john_day_river_bon_adult_returns_2005-2015.csv")))
```

#### Inspect data
```{r}
# Unique tags?
length(unique(JDR_BON_adult_returns$tag_code))
# 2,121 unique tags - and they're all wild (or unknown)! This matches Shelby's numbers for wild fish.

# Count by year
# Reformat dates
JDR_BON_adult_returns %>% 
  mutate(obs_time_value = mdy_hms(obs_time_value)) %>% 
  mutate(year = year(obs_time_value)) -> JDR_BON_adult_returns
```

#### Export tags for querying PTAGIS
```{r}
as.data.frame(unique(JDR_BON_adult_returns$tag_code)) -> JDR_unique_tags

write.table(JDR_unique_tags, file = here(here("PTAGIS_queries", "intermediate_files",  "2022-01-06-john_day_river_bon_adult_returns_2005-2015_TAGS.txt")), col.names = FALSE, row.names = FALSE, quote = FALSE)
```

## QUERY 2
- Complete tag history

Query Builder2:
- 1 Attributes: Standard plus event site metadata fields: Event Site Basin, Event Site Info, Event Site Code, Event Site Subbasin, Event Site RKM
- 2 Metrics: Standard
- 29 Tag Code: 2022-01-06-john_day_river_bon_adult_returns_2005-2015_TAGS.txt

# Analysis

## Import and clean data
```{r}
JDR_CTH <- clean_names(read.csv(here("PTAGIS_queries", "complete_tag_histories", "2022-01-07-jdr_2005_2015_complete_tag_history.csv")))
JDR_CTH %>% 
  mutate(event_date = mdy(event_date_mmddyyyy)) %>% 
  arrange(tag_code, event_date) %>% 
  dplyr::select(tag_code, event_type_name, event_site_name, event_site_basin_name, event_site_subbasin_name, event_site_rkm_value, cth_count, event_date) -> JDR_CTH

JDR_CTH
```


Determine run timing based on the date/time of first observation at:
- Bonneville Dam
- overshoot dams (first dam with PIT detection capabilities upstream of the natal tributary)
  - Look in adult latter, or time of first observation above overshoot dam
- in natal tributary

Sort into run years by June 1

## Turn each detection history into characterization of run timing

### Make a df to sort by run year
```{r}
run_year <- c("05/06", "06/07", "07/08", "08/09", "09/10", "10/11", "11/12", "12/13", "13/14", "14/15")
run_year_start <- seq(ymd("2005-06-01"), ymd("2014-06-01"), by = "years")
run_year_end <- seq(ymd("2006-05-31"), ymd("2015-05-31"), by = "years")

run_year_df <- data.frame(run_year, run_year_start, run_year_end)

subset(run_year_df, run_year_start < BON_arrival_date & run_year_end > BON_arrival_date)$run_year
```



### Determine tally-based overshoot, homing, fallback rates
```{r detection_history_to_classification}
# Get a vector of all unique tags
unique_tags <- unique(JDR_CTH$tag_code)

# Figure out all the places where fish were detected
as.data.frame(table(JDR_CTH$event_site_name))

# Make a dataframe of the sites and their metadata
JDR_CTH[!duplicated(JDR_CTH$event_site_name),] %>% 
  dplyr::select(-c(tag_code, event_type_name, cth_count, event_date)) -> site_metadata
  
  

# Get the names of all Bonneville adult fishway detectors
BON_adult_fishways <- c("BHL - Adult Fishway at BONH", "BO1 - Bonneville Bradford Is. Ladder", "BO2 - Bonneville Cascades Is. Ladder", "BO3 - Bonneville WA Shore Ladder/AFF", "BO4 - Bonneville WA Ladder Slots", "BONAFF - BON - Adult Fish Facility")

# Get the names of all in stream arrays in John Day River
JDR_instream_arrays <- subset(site_metadata, event_site_basin_name == "John Day")$event_site_name

# Get the names of all arrays that would indicate overshoot (primarily McNary dam, but also other upstream arrays)
overshoot_arrays <- subset(site_metadata, event_site_basin_name %in% c("Lower Snake", "Clearwater", "Salmon", "Upper Columbia", "Yakima")  | event_site_subbasin_name %in% c("Middle Columbia-Lake Wallula", "Umatilla", "Walla Walla" ))$event_site_name

# Make a dataframe to store the fish status
overshoot_df <- data.frame(tag = unique_tags, run_year = NA, overshoot = NA, home = NA, fallback = NA)

# For each of the unique tags
for (i in 1:length(unique_tags)){
  tag_ID <- unique_tags[i]
  tag_hist <- subset(JDR_CTH, tag_code == tag_ID)
  
  # Get the date of arrival at BON - take min in case it fell back over Bonneville
  BON_arrival_date <- min(subset(tag_hist, event_site_name %in% c("BHL - Adult Fishway at BONH", "BO1 - Bonneville Bradford Is. Ladder", "BO2 - Bonneville Cascades Is. Ladder", "BO3 - Bonneville WA Shore Ladder/AFF", "BO4 - Bonneville WA Ladder Slots", "BONAFF - BON - Adult Fish Facility"))$event_date)
  
  # Assign the fish to the correct run year
  overshoot_df[i,2] <- subset(run_year_df, run_year_start < BON_arrival_date & run_year_end > BON_arrival_date)$run_year
  
  # Arrange by date
  tag_hist %>% 
    arrange(event_date) %>% 
    # Remove all detections that occur before detection at Bonneville upstream
    subset(., event_date >= BON_arrival_date) -> tag_hist_adult
  
  # Determine if overshot
  tag_hist_adult$event_site_name %in% overshoot_arrays -> overshoot_TF
  if (TRUE %in% overshoot_TF){
    overshoot_df[i,3] <- TRUE
  }
  else{
    overshoot_df[i,3] <- FALSE
  }
  
  # If it made it home, home = TRUE
  tag_hist_adult$event_site_name %in% JDR_instream_arrays -> instream_TF
  if (TRUE %in% instream_TF){
    overshoot_df[i,4] <- TRUE
  }
  else{
    overshoot_df[i,4] <- FALSE
  }
}
```
## Examine rates

### Total success rate
```{r}
table(overshoot_df$home)

# Tally estimate suggests only 32.39% successfully homed. But that is a big underestimate because detection efficiency in the natal tributary is far from 100%.
# According to Shelby's estimate, tally underestimates successful homing by about 8%
# Table A.2 (PDF page 154) - detection efficiency at JD1 mean of 71%, but is around 99% from 08-09 to 10/11, then drops to 40ish% from 12/13 to 14/15.
# Note in this table: “JD1” was washed out and replaced in 2011/2012

# Look into the detection histories of those that failed to home
failed_home_tags <- subset(overshoot_df, home == FALSE)$tag

subset(JDR_CTH, tag_code %in% failed_home_tags) %>% 
  dplyr::select(tag_code, event_site_name, event_site_subbasin_name, event_date) 


table(overshoot_df$run_year)
# The numbers are slightly different than Shelby's - I assume it's because of her manual editing of run years. From her thesis: 
# "Next, observations were sorted into run years. Because most populations of interior Columbia River basin steelhead normally begin their upstream migration in summer and spawn the following spring (Robards and Quinn 2002), 1 June was used as the division between run years. No date would be a perfect division between runs for all fish in all tributaries. To account for this, I manually examined the individual capture histories of all fish that had observations in consecutive run years and assigned fish to the correct run year or years. Some fish had single migrations that overlapped run years, with either early detections at downstream sites, or late detections indicating downstream movement after the spawning season. Other fish made multiple consecutive upstream migrations (successful kelts that returned as repeat spawners) and were kept in both run years."

# Look at the table of success rates by run year plus N
overshoot_df %>% 
  group_by(run_year) %>% 
  summarise(home_rate = sum(home), n = n()) %>% 
  mutate(home_rate = home_rate/n) -> total_success_tally_by_year

total_success_tally_by_year
# This looks very similar to Shelby's Figure A.3, for the years prior to 2012
# 2011/2012 is a wash because the array was down
```


### Overshoot
```{r}
table(overshoot_df$overshoot)

# Look at the table of success rates by run year plus N
overshoot_df %>% 
  group_by(run_year) %>% 
  summarise(overshoot_rate = sum(overshoot), n = n()) %>% 
  mutate(overshoot_rate = overshoot_rate/n) -> overshoot_tally_by_year

overshoot_tally_by_year
# These again look very similar to Shelby's results - Page 44, Fig 2.8
```

### Fallback to home
```{r}
overshoot_df %>% 
  mutate(fallback = ifelse(overshoot == TRUE & home == TRUE, TRUE,
                           ifelse(overshoot == TRUE & home == FALSE, FALSE, NA))) -> overshoot_df

# Look at the table of success rates by run year plus N
overshoot_df %>% 
  subset(., !is.na(fallback)) %>% 
  group_by(run_year) %>% 
  summarise(fallback_rate = sum(fallback), n = n()) %>% 
  mutate(fallback_rate = fallback_rate/n) -> fallback_tally_by_year

fallback_tally_by_year
# Looks similar to Shelby's Fig 2.11 (Page 38) - I think she only has estimates for 08/09, 09/10, 10/11, and then 12/13, 13/14, and 14/15 because of sample sizes
```



# Project Branch
DOESN'T WORK ON MacOS! Time to manually code these models...


# Multistate Release-Recapture Model

Let's see if we can write it in a maximum likelihood framework prior to going into JAGS/Bayesian.

For pre-2017 John Day River Steelhead:

Probability of being seen overshooting at MCN (for fish seen at Bonneville):

prob(moving from BON -> TDA) * prob(detection at TDA) * prob(moving from TDA -> MCN) * prob(detection at MCN)

"Models of John Day and Umatilla populations did not include early migration success, because the first dam with adult detectors below the home tributary was Bonneville Dam." (Richins 2017)
- So it sounds like we should ignore the Dalles, because apparently the adult detectors there didn't really work?

See Table E.1 (page 198): conditional overshoot rates
- Calculated for BON -> MCN, MCN -> ICH, ICH -> LGR, and MCN -> PRA
- TDA isn't mentioned

For other tables of conditional overshoot rates, it's the dam below the natal tributary to the dams upstream. So we're only looking at the first dam below natal tributary with adult detectors, which apparently doesn't include TDA.

## Get data into format for multinomial likelihood
```{r reformat_data}

JDR_CTH[!duplicated(JDR_CTH$event_site_name),] %>% 
  dplyr::select(-c(tag_code, event_type_name, cth_count, event_date)) -> site_metadata
  

# Get the names of all Bonneville adult fishway detectors
BON_adult_fishways <- c("BHL - Adult Fishway at BONH", "BO1 - Bonneville Bradford Is. Ladder", "BO2 - Bonneville Cascades Is. Ladder", "BO3 - Bonneville WA Shore Ladder/AFF", "BO4 - Bonneville WA Ladder Slots", "BONAFF - BON - Adult Fish Facility")

# Get names of all The Dalles adult fishway detectors
site_metadata$event_site_name[grep("Dalles", site_metadata$event_site_name)]
TDA_adult_fishways <- c("TD1 - The Dalles East Fish Ladder", "TD2 - The Dalles North Fish Ladder")

# Get names of all McNary adult fishway detectors
site_metadata$event_site_name[grep("McNary", site_metadata$event_site_name)]
MCN_adult_fishways <- c("MC1 - McNary Oregon Shore Ladder",  "MC2 - McNary Washington Shore Ladder")

# Get names of all ICH adult fishway detectors
site_metadata$event_site_name[grep("Ice", site_metadata$event_site_name)]
ICH_adult_fishways <- c("ICH - Ice Harbor Dam (Combined)")

# Get names of all LGR adult fishway detectors
site_metadata$event_site_name[grep("Granite", site_metadata$event_site_name)]
LGR_adult_fishways <- c("GRA - Lower Granite Dam Adult")

# Get names of all PRA adult fishway detectors
site_metadata$event_site_name[grep("Priest", site_metadata$event_site_name)]
PRA_adult_fishways <- c("PRA - Priest Rapids Adult")

# Get the names of all in stream arrays in John Day River
JDR_instream_arrays <- subset(site_metadata, event_site_basin_name == "John Day")$event_site_name

# Get the names of all arrays that would indicate overshoot (primarily McNary dam, but also other upstream arrays)
overshoot_arrays <- subset(site_metadata, event_site_basin_name %in% c("Lower Snake", "Clearwater", "Salmon", "Upper Columbia", "Yakima")  | event_site_subbasin_name %in% c("Middle Columbia-Lake Wallula", "Umatilla", "Walla Walla" ))$event_site_name

# Make a dataframe to store fish detection history
detection_df <- data.frame(tag = unique_tags, run_year = NA, BON = NA, TDA = NA, MCN = NA, ICH = NA, LGR = NA, PRA = NA, natal = NA)

# Record that they were seen at BON
detection_df$BON <- TRUE

# For each of the unique tags
for (i in 1:length(unique_tags)){
  tag_ID <- unique_tags[i]
  tag_hist <- subset(JDR_CTH, tag_code == tag_ID)
  
  # Get the date of arrival at BON - take min in case it fell back over Bonneville
  BON_arrival_date <- min(subset(tag_hist, event_site_name %in% c("BHL - Adult Fishway at BONH", "BO1 - Bonneville Bradford Is. Ladder", "BO2 - Bonneville Cascades Is. Ladder", "BO3 - Bonneville WA Shore Ladder/AFF", "BO4 - Bonneville WA Ladder Slots", "BONAFF - BON - Adult Fish Facility"))$event_date)
  
  # Assign the fish to the correct run year
  detection_df[i,2] <- subset(run_year_df, run_year_start < BON_arrival_date & run_year_end > BON_arrival_date)$run_year
  
  # Arrange by date
  tag_hist %>% 
    arrange(event_date) %>% 
    # Remove all detections that occur before detection at Bonneville upstream
    subset(., event_date >= BON_arrival_date) -> tag_hist_adult
  
  
  # Record if seen at TDA
  tag_hist_adult$event_site_name %in% TDA_adult_fishways -> TDA_TF
  if (TRUE %in% TDA_TF){
    detection_df[i,4] <- TRUE
  }
  else{
    detection_df[i,4] <- FALSE
  }
  
  # Record if seen at MCN
  tag_hist_adult$event_site_name %in% MCN_adult_fishways -> MCN_TF
  if (TRUE %in% MCN_TF){
    detection_df[i,5] <- TRUE
  }
  else{
    detection_df[i,5] <- FALSE
  }
    
  # Record if seen at ICH
  tag_hist_adult$event_site_name %in% ICH_adult_fishways -> ICH_TF
  if (TRUE %in% ICH_TF){
    detection_df[i,6] <- TRUE
  }
  else{
    detection_df[i,6] <- FALSE
  }
  
  # Record if seen at LGR
  tag_hist_adult$event_site_name %in% LGR_adult_fishways -> LGR_TF
  if (TRUE %in% LGR_TF){
    detection_df[i,7] <- TRUE
  }
  else{
    detection_df[i,7] <- FALSE
  }
  
  # Record if seen at PRA
  tag_hist_adult$event_site_name %in% PRA_adult_fishways -> PRA_TF
  if (TRUE %in% PRA_TF){
    detection_df[i,8] <- TRUE
  }
  else{
    detection_df[i,8] <- FALSE
  }
  
  # If it made it home, natal = TRUE
  tag_hist_adult$event_site_name %in% JDR_instream_arrays -> instream_TF
  if (TRUE %in% instream_TF){
    detection_df[i,9] <- TRUE
  }
  else{
    detection_df[i,9] <- FALSE
  }
}

detection_df

# Summarise into counts by site
detection_df %>% 
  summarise_at(vars(BON:natal), sum) -> detection_counts_v1

detection_counts_v1
# Doesn't work - need transitions, not counts at each sites

# as.data.frame(table(JDR_CTH$event_site_name)) %>% 
#   arrange(desc(Freq))
```

## Get data into format for multinomial likelihood
```{r reformat_data}

JDR_CTH[!duplicated(JDR_CTH$event_site_name),] %>% 
  dplyr::select(-c(tag_code, event_type_name, cth_count, event_date)) -> site_metadata
  

# Get the names of all Bonneville adult fishway detectors
BON_adult_fishways <- c("BHL - Adult Fishway at BONH", "BO1 - Bonneville Bradford Is. Ladder", "BO2 - Bonneville Cascades Is. Ladder", "BO3 - Bonneville WA Shore Ladder/AFF", "BO4 - Bonneville WA Ladder Slots", "BONAFF - BON - Adult Fish Facility")

# Get names of all The Dalles adult fishway detectors
site_metadata$event_site_name[grep("Dalles", site_metadata$event_site_name)]
TDA_adult_fishways <- c("TD1 - The Dalles East Fish Ladder", "TD2 - The Dalles North Fish Ladder")

# Get names of all McNary adult fishway detectors
site_metadata$event_site_name[grep("McNary", site_metadata$event_site_name)]
MCN_adult_fishways <- c("MC1 - McNary Oregon Shore Ladder",  "MC2 - McNary Washington Shore Ladder")

# Get names of all ICH adult fishway detectors
site_metadata$event_site_name[grep("Ice", site_metadata$event_site_name)]
ICH_adult_fishways <- c("ICH - Ice Harbor Dam (Combined)")

# Get names of all LGR adult fishway detectors
site_metadata$event_site_name[grep("Granite", site_metadata$event_site_name)]
LGR_adult_fishways <- c("GRA - Lower Granite Dam Adult")

# Get names of all PRA adult fishway detectors
site_metadata$event_site_name[grep("Priest", site_metadata$event_site_name)]
PRA_adult_fishways <- c("PRA - Priest Rapids Adult")

# Get the names of all in stream arrays in John Day River
JDR_instream_arrays <- subset(site_metadata, event_site_basin_name == "John Day")$event_site_name

# Get the names of all arrays that would indicate overshoot (primarily McNary dam, but also other upstream arrays)
overshoot_arrays <- subset(site_metadata, event_site_basin_name %in% c("Lower Snake", "Clearwater", "Salmon", "Upper Columbia", "Yakima")  | event_site_subbasin_name %in% c("Middle Columbia-Lake Wallula", "Umatilla", "Walla Walla" ))$event_site_name

# Make a dataframe to store fish detection history
detection_df <- data.frame(tag = unique_tags, run_year = NA, BON = NA, BON_TDA = NA, BON_MCN = NA, MCN_ICH = NA, ICH_LGR = NA, MCN_PRA = NA, natal = NA)

# Record that they were seen at BON
detection_df$BON <- TRUE

# For each of the unique tags
for (i in 1:length(unique_tags)){
  tag_ID <- unique_tags[i]
  tag_hist <- subset(JDR_CTH, tag_code == tag_ID)
  
  # Get the date of arrival at BON - take min in case it fell back over Bonneville
  BON_arrival_date <- min(subset(tag_hist, event_site_name %in% c("BHL - Adult Fishway at BONH", "BO1 - Bonneville Bradford Is. Ladder", "BO2 - Bonneville Cascades Is. Ladder", "BO3 - Bonneville WA Shore Ladder/AFF", "BO4 - Bonneville WA Ladder Slots", "BONAFF - BON - Adult Fish Facility"))$event_date)
  
  # Assign the fish to the correct run year
  detection_df[i,2] <- subset(run_year_df, run_year_start < BON_arrival_date & run_year_end > BON_arrival_date)$run_year
  
  # Arrange by date
  tag_hist %>% 
    arrange(event_date) %>% 
    # Remove all detections that occur before detection at Bonneville upstream
    subset(., event_date >= BON_arrival_date) -> tag_hist_adult
  
  
  # Record if seen at BON & TDA (BON -> TDA)
  tag_hist_adult$event_site_name %in% TDA_adult_fishways -> TDA_TF
  if (TRUE %in% TDA_TF & detection_df[i,3] == TRUE){
    detection_df[i,4] <- TRUE
  }
  else{
    detection_df[i,4] <- FALSE
  }
  
  # Record if seen at BON & MCN (BON -> MCN)
  tag_hist_adult$event_site_name %in% MCN_adult_fishways -> MCN_TF
  if (TRUE %in% MCN_TF  & detection_df[i,3] == TRUE){
    detection_df[i,5] <- TRUE
  }
  else{
    detection_df[i,5] <- FALSE
  }
    
  # Record if seen at ICH & MCN (MCN -> ICH)
  tag_hist_adult$event_site_name %in% ICH_adult_fishways -> ICH_TF
  if (TRUE %in% MCN_TF & TRUE %in% ICH_TF){
    detection_df[i,6] <- TRUE
  }
  else{
    detection_df[i,6] <- FALSE
  }
  
  # Record if seen at ICH & LGR
  tag_hist_adult$event_site_name %in% LGR_adult_fishways -> LGR_TF
  if (TRUE %in% LGR_TF & TRUE %in% ICH_TF){
    detection_df[i,7] <- TRUE
  }
  else{
    detection_df[i,7] <- FALSE
  }
  
  # Record if seen at MCN & PRA
  tag_hist_adult$event_site_name %in% PRA_adult_fishways -> PRA_TF
  if (TRUE %in% PRA_TF & TRUE %in% MCN_TF){
    detection_df[i,8] <- TRUE
  }
  else{
    detection_df[i,8] <- FALSE
  }
  
  # If it made it home, natal = TRUE
  tag_hist_adult$event_site_name %in% JDR_instream_arrays -> instream_TF
  if (TRUE %in% instream_TF){
    detection_df[i,9] <- TRUE
  }
  else{
    detection_df[i,9] <- FALSE
  }
}

detection_df

# Summarise into counts by site
detection_df %>% 
  summarise_at(vars(BON:natal), sum) -> detection_counts

detection_counts

# Get rid of BON_TDA
detection_counts %>% 
  dplyr::select(-BON_TDA) -> detection_counts
```


### Writing the multinomial likelihood:

The parameters that we are estimating:

Migration successes (aka survival):
- BON_MCN
- MCN_ICH
- ICH_LGR
- MCN_PRA
- BON_natal
- overshoot_natal

Detection efficiency:
- MCN
- ICH
- LGR
- PRA
- natal

Choice of movement direction (branching)
- After Bonneville, they can either home directly (would be seen in natal tributary) OR they can overshoot MCN

Need to calculate overshoot rates (conditional and unconditional).

Richins 2017: Unconditional overshooting rates were calculated as the proportion of fish seen in the fish ladder of the overshot dam out of the number of fish seen at Bonneville, divided by the detection efficiency at the overshot dam in that run year (Eq. 2.2).

This is a branching system (not linear).

Let's do a simple system - just a linear system, looking at overshoot only at MCN.
- parameters:
  - survival BON -> MCN
  - survival BON -> natal
  - branch after BON; can overshoot (MCN) or home directly (natal)
  - detection efficiency at MCN
  - detection efficiency at natal

But I think that you need the upstream detectors too in order to estimate detection efficiency?

```{r multinomial_likelihood}
x <- detection_counts[, c(1,2,3)]
p <- c(1,0.53, 0.25)

dmultinom(x, prob = p)

p.MLE = function(n, y){
  negLL = function(p) (-1)*dbinom(y, size = n, prob = p, log = TRUE)
  fit = optim(0.2, negLL, method = 'L-BFGS-B', hessian = TRUE, lower = 0.0001, upper = 0.9999)
  var = solve(fit$hessian)
  list(logLik = fit$value, MLE = fit$par, var = var)
}
p.MLE(n = 20, y = 12)


dmultinom()

```



