---
title: "covariate_windows"
author: "Markus Min"
date: '2023-04-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# load libraries
library(here)
library(janitor)
library(tidyverse)
library(lubridate)
library(car)
library(ggthemes)

# load data


```

```{r load_states_data}
# so, I think that the most up to date files are in the same folders as the model runs
snake_adults_states_complete <- read.csv(here::here("stan_actual", "deteff_ESU_models", "500iter_3chain_runs", "snake", "snake_adults_states_complete.csv"), row.names = 1)
lowcol_adults_states_complete <- read.csv(here::here("stan_actual", "deteff_ESU_models", "lower_columbia", "lower_columbia_adults_states_complete.csv"), row.names = 1)
midcol_adults_states_complete <- read.csv(here::here("stan_actual", "deteff_ESU_models", "500iter_3chain_runs", "middle_columbia", "middle_columbia_adults_states_complete.csv"), row.names = 1)
uppcol_adults_states_complete <- read.csv(here::here("stan_actual", "deteff_ESU_models", "500iter_3chain_runs", "upper_columbia", "upper_columbia_adults_states_complete.csv"), row.names = 1)

# combine them all
snake_adults_states_complete %>% 
  bind_rows(., midcol_adults_states_complete) %>% 
  bind_rows(., uppcol_adults_states_complete) %>% 
  bind_rows(., lowcol_adults_states_complete) -> ASC

# now add tag code metadata, for natal origins
origin_numeric <- data.frame(natal_origin = c("Asotin_Creek", 
                                        "Clearwater_River",
                                        "Deschutes_River", 
                                        "Entiat_River", 
                                        "Fifteenmile_Creek", 
                                        "Grande_Ronde_River", 
                                        "Hood_River",
                                        "Imnaha_River",
                                        "John_Day_River", 
                                        "Methow_River", 
                                        "Okanogan_River", 
                                        "Salmon_River", 
                                        "Tucannon_River", 
                                        "Umatilla_River",
                                        "Walla_Walla_River",
                                        "Wenatchee_River", 
                                        "Yakima_River"),
                             natal_origin_numeric = seq(1,17,1))

origin_rear_actual <- read.csv(here::here("stan_actual", "origin_rear_actual.csv"), row.names = 1)
origin_rear_actual %>% 
  dplyr::rename(natal_origin_numeric = natal_origin) %>% 
  left_join(origin_numeric, by = "natal_origin_numeric") %>% 
  dplyr::select(tag_code_2, natal_origin) -> tag_code_origins

ASC %>% 
  left_join(., tag_code_origins, by = "tag_code_2") -> ASC

# also note which ESU they're from, for plotting
ESU_origins <- data.frame(natal_origin = unique(ASC$natal_origin),
                          ESU = c(rep("Snake", 6),
                                  rep("Middle Columbia", 6),
                                  rep("Upper Columbia", 4),
                                  "Lower Columbia"))

ASC %>% 
  left_join(., ESU_origins, by = "natal_origin") -> ASC
```

```{r functions}
# a function that takes two states, and finds the average amount of time between those states
# notably, if a fish doesn't go straight, it won't keep it in this function
# so, if a fish attempts BON ascent on day 1, gives up, and does it again on day 7, then is seen at MCN on day 17,
# passage time will be 10 days, not 16 days
# also, if a fish ascends BON, then goes to Deschutes, then goes to MCN, it will not be kept in this function

passage_time <- function(pathway_1, pathway_2){
  
  # first, subset only those that have those two states
  # don't keep any with implicit time interpolated
  ASC %>% 
    filter(pathway == pathway_1 & lead(pathway == pathway_2) & tag_code_2 == lead(tag_code_2) & pathway != "implicit" & lead(pathway) != "implicit" |
             pathway == pathway_2 & lag(pathway == pathway_1) & tag_code_2 == lag(tag_code_2) & pathway != "implicit" & lag(pathway) != "implicit") %>% 
    mutate(date_time = ymd_hms(date_time)) -> two_pathway_df
  
  # make a data frame to record all of the transitions
  n_transitions <- nrow(two_pathway_df)/2
  passage_df <- data.frame(tag_code_2 = two_pathway_df$tag_code_2[seq(1,(nrow(two_pathway_df)-1),2)],
                           ESU = two_pathway_df$ESU[seq(1,(nrow(two_pathway_df)-1),2)],
                           passage_time = NA)
  
  
  for (i in 1:nrow(passage_df)){
    passage_df$passage_time[i] <- two_pathway_df$date_time[(i*2)] - two_pathway_df$date_time[(i*2-1)]
    
  }
  
  return(passage_df)
  
}

# a function to plot passage time
passage_time_plot <- function(passage_df, reach){
  passage_plot <- ggplot(passage_df, aes(x = passage_time, fill = ESU)) +
    geom_histogram(binwidth = 1) +
    xlab("Days") +
    ggtitle(reach) +
    scale_fill_tableau(palette = "Tableau 10")
 
  return(passage_plot) 
}

passage_time_plot_quantiles <- function(passage_df, reach, quantiles){
  passage_plot <- ggplot(passage_df, aes(x = passage_time, fill = ESU)) +
    geom_histogram(binwidth = 1) +
    xlab("Days") +
    ggtitle(reach) +
    scale_fill_tableau(palette = "Tableau 10") +
    geom_vline(xintercept = quantiles, lty = 2)
 
  return(passage_plot) 
}

passage_time_plot_logdays <- function(passage_df, reach){
  passage_plot <- ggplot(passage_df, aes(x = log(passage_time), fill = ESU)) +
    geom_histogram(binwidth = 0.1) +
    xlab("log(Days)") +
    ggtitle(reach) +
    scale_fill_tableau(palette = "Tableau 10")
  
  return(passage_plot) 
}

passage_time_plot_logdays_quantiles <- function(passage_df, reach, quantiles){
  passage_plot <- ggplot(passage_df, aes(x = log(passage_time), fill = ESU)) +
    geom_histogram(binwidth = 0.1) +
    xlab("log(Days)") +
    ggtitle(reach) +
    scale_fill_tableau(palette = "Tableau 10") +
    geom_vline(xintercept = quantiles, lty = 2)
  
  return(passage_plot) 
}

passage_time_plot_origin <- function(passage_df, reach){
  passage_plot <- ggplot(passage_df, aes(x = passage_time, fill = natal_origin)) +
    geom_histogram(binwidth = 1) +
    xlab("Days") +
    ggtitle(reach) +
    scale_fill_tableau(palette = "Tableau 20")
 
  return(passage_plot) 
}

passage_time_plot_logdays_origin <- function(passage_df, reach){
  passage_plot <- ggplot(passage_df, aes(x = log(passage_time), fill = natal_origin)) +
    geom_histogram(binwidth = 0.1) +
    xlab("log(Days)") +
    ggtitle(reach) +
    scale_fill_tableau(palette = "Tableau 20")
  
  return(passage_plot) 
}


# a function that takes just one state, and finds the average amount of time to move out of that state

residence_time <- function(residence_state){
  
  # don't keep any with implicit time interpolated
  ASC %>% 
    filter(state == residence_state & tag_code_2 == lead(tag_code_2) & pathway != "implicit" & lead(pathway) != "implicit" |
             lag(state) == residence_state & tag_code_2 == lag(tag_code_2) & pathway != "implicit" & lag(pathway) != "implicit") %>% 
    mutate(date_time = ymd_hms(date_time)) -> one_state_df
  
  # make a data frame to record all of the transitions
  n_transitions <- nrow(one_state_df)/2
  passage_df <- data.frame(tag_code_2 = one_state_df$tag_code_2[seq(1,(nrow(one_state_df)-1),2)],
                           ESU = one_state_df$ESU[seq(1,(nrow(one_state_df)-1),2)],
                           natal_origin = one_state_df$natal_origin[seq(1,(nrow(one_state_df)-1),2)],
                           passage_time = NA)
  
  
  for (i in 1:nrow(passage_df)){
    passage_df$passage_time[i] <- one_state_df$date_time[(i*2)] - one_state_df$date_time[(i*2-1)]
    
  }
  
  return(passage_df)
  
}
```

```{r BON_to_MCN, eval = FALSE}

BON_MCN_passage <- passage_time(pathway_1 = "BON (adult)", pathway_2 = "MCN (adult)")
BON_to_MCN <- passage_time_plot(passage_df = BON_MCN_passage, reach = "Bonneville Adult Ladder to McNary Adult Ladder")
BON_to_MCN

BON_to_MCN_log <- passage_time_plot_logdays(passage_df = BON_MCN_passage, reach = "Bonneville Adult Ladder to McNary Adult Ladder")
BON_to_MCN_log

# give summary of passage time

summary(BON_MCN_passage$passage_time)
```

```{r all_states_residence_time}
unique_states <- unique(ASC$state)
# okay, so some of these are terminal states, where literally no fish that go in come out
# Salmon river
unique_states <- unique_states[!(unique_states == "Salmon River")]
unique_states <- unique_states[!(unique_states == "Clearwater River")]
unique_states <- unique_states[!(unique_states == "loss")]
unique_states <- unique_states[!(unique_states == "Upstream WEL other tributaries")]

# we might be interested in just mainstem sites, but for now we'll keep in tribs
mainstem_states <- c("mainstem, mouth to BON",
  "mainstem, BON to MCN",
  "mainstem, MCN to ICH or PRA",
  "mainstem, PRA to RIS",
  "mainstem, RIS to RRE",
  "mainstem, RRE to WEL",
  "mainstem, upstream of WEL",
  "mainstem, ICH to LGR",
  "mainstem, upstream of LGR")

plots_days <- list()
plots_logdays <- list()
plots_days_origin <- list()
plots_logdays_origin <- list()

for (i in 1:length(mainstem_states)) {
  residence_df <- residence_time(residence_state = mainstem_states[i])
  
  passage_plot <- passage_time_plot(passage_df = residence_df, reach = mainstem_states[i])
  plots_days[[i]] <- passage_plot
  
  passage_plot_origin <- passage_time_plot_origin(passage_df = residence_df, reach = mainstem_states[i])
  plots_days_origin[[i]] <- passage_plot_origin
  
  passage_plot_log <- passage_time_plot_logdays(passage_df = residence_df, reach = mainstem_states[i])
  plots_logdays[[i]] <- passage_plot_log
  
  passage_plot_log_origin <- passage_time_plot_logdays_origin(passage_df = residence_df, reach = mainstem_states[i])
  plots_logdays_origin[[i]] <- passage_plot_log_origin

  # give summary of passage time
  summary(residence_df$passage_time)
  
}

```


```{r all_states_residence_time_window_exploration}
main_mainstem_states <- c(
  "mainstem, BON to MCN",
  "mainstem, MCN to ICH or PRA",
  "mainstem, PRA to RIS",
  "mainstem, RIS to RRE",
  "mainstem, RRE to WEL",
  "mainstem, ICH to LGR")

plots_days_windows <- list()
plots_logdays_windows <- list()
plots_days_origin_windows <- list()
plots_logdays_origin_windows <- list()
list_quantiles <- list()

for (i in 1:length(main_mainstem_states)) {
  residence_df <- residence_time(residence_state = main_mainstem_states[i])
  
  # Look into windows - show: 1) median of the whole distribution, 
  # and what cutoff you'd need to capture 50%, 80%, and 95% of fish residence time
  quantiles <- quantile(residence_df$passage_time, probs = c(0.5, 0.8, 0.95))
  log(quantiles)
  list_quantiles[[i]] <- quantiles
  
  passage_plot_quantiles <- passage_time_plot_quantiles(passage_df = residence_df, reach = main_mainstem_states[i], quantiles = quantiles)
  plots_days_windows[[i]] <- passage_plot_quantiles
  
  passage_plot_log_quantiles <- passage_time_plot_logdays_quantiles(passage_df = residence_df, reach = main_mainstem_states[i], quantiles = log(quantiles))
  plots_logdays_windows[[i]] <- passage_plot_log_quantiles
  
}

```

## Residence time by main mainstem reaches

This table shows for each mainstem reach, what the median (t50), 80th quantile (t80), and 95th quantile (t95) residence time is for all fish that passed through this state.

The dotted lines in the following histograms show where these quantiles fall.
```{r}

# make an actual table
quantiles_table <- data.frame(reach = main_mainstem_states,
                              t50 = c(list_quantiles[[1]][1],
                                       list_quantiles[[2]][1],
                                       list_quantiles[[3]][1],
                                       list_quantiles[[4]][1],
                                       list_quantiles[[5]][1],
                                       list_quantiles[[6]][1]),
                              t80 = c(list_quantiles[[1]][2],
                                       list_quantiles[[2]][2],
                                       list_quantiles[[3]][2],
                                       list_quantiles[[4]][2],
                                       list_quantiles[[5]][2],
                                       list_quantiles[[6]][2]),
                              t95 = c(list_quantiles[[1]][3],
                                       list_quantiles[[2]][3],
                                       list_quantiles[[3]][3],
                                       list_quantiles[[4]][3],
                                       list_quantiles[[5]][3],
                                       list_quantiles[[6]][3]))

write.csv(quantiles_table, here::here("covariate_data", "reach_quantiles.csv"))

plots_days_windows[[1]]
# list_quantiles[[1]]
plots_logdays_windows[[1]]
plots_days_windows[[2]]
# list_quantiles[[2]]
plots_logdays_windows[[2]]
plots_days_windows[[3]]
# list_quantiles[[3]]
plots_logdays_windows[[3]]
plots_days_windows[[4]]
# list_quantiles[[4]]
plots_logdays_windows[[4]]
plots_days_windows[[5]]
# list_quantiles[[5]]
plots_logdays_windows[[5]]
plots_days_windows[[6]]
# list_quantiles[[6]]
plots_logdays_windows[[6]]

```

### how correlated is temperature in these different windows? 

NOTE: here, seasons are defined as:

- Winter = Jan/Feb/March
- Spring = April/May/June
- Summer = July/August/September
- Fall = October/November/December

#### McNary Dam - using the BON to MCN residence times, which have the widest spread
```{r}
MCN_temp_long <- read.csv(here::here("covariate_data", "for_model", "MCN_temp.csv"))

MCN_t50 <- round(subset(quantiles_table, reach == "mainstem, BON to MCN")$t50, 0)
MCN_t80 <- round(subset(quantiles_table, reach == "mainstem, BON to MCN")$t80, 0)
MCN_t95 <- round(subset(quantiles_table, reach == "mainstem, BON to MCN")$t95, 0)

# first, drop the NAs
MCN_temp_long %>% 
  filter(!is.na(temp)) %>% 
  mutate(date = date(date)) %>% 
  # create a season field
  mutate(month = month(date)) %>% 
  mutate(season = ifelse(month %in% c(1,2,3), "Winter",
                         ifelse(month %in% c(4,5,6), "Spring",
                                ifelse(month %in% c(7,8,9), "Summer",
                                       ifelse(month %in% c(10,11,12), "Fall",
                                              "ERROR"))))) %>% 
  mutate(t50 = NA, t80 = NA, t95 = NA) %>% 
  # let's start in 2005, since that's when our data starts
  subset(year >= 2005) -> MCN_temp_long

# loop to get all windows
for (i in 1:nrow(subset(MCN_temp_long, year <= 2022))){
  MCN_temp_long$t50[i] <- mean(subset(MCN_temp_long, date >= MCN_temp_long$date[i] & date <= MCN_temp_long$date[i]+ days(x = MCN_t50))$temp)
  MCN_temp_long$t80[i] <- mean(subset(MCN_temp_long, date >= MCN_temp_long$date[i] & date <= MCN_temp_long$date[i]+ days(x = MCN_t80))$temp)
  MCN_temp_long$t95[i] <- mean(subset(MCN_temp_long, date >= MCN_temp_long$date[i] & MCN_temp_long$date[i]+ days(x = MCN_t95))$temp)
}

# plot the correlation
ggplot(MCN_temp_long, aes(x = t50, y = t80, color = season)) +
  geom_point() +
  scale_color_tableau(palette = "Tableau 10")

ggplot(MCN_temp_long, aes(x = t50, y = t95, color = season)) +
  geom_point() +
  scale_color_tableau(palette = "Tableau 10")
```

#### Ice Harbor Dam - using the MCN to ICH or PRA residence times, which have a much narrower spread
```{r}
ICH_temp_long <- read.csv(here::here("covariate_data", "for_model", "ICH_temp.csv"))

ICH_t50 <- round(subset(quantiles_table, reach == "mainstem, MCN to ICH or PRA")$t50, 0)
ICH_t80 <- round(subset(quantiles_table, reach == "mainstem, MCN to ICH or PRA")$t80, 0)
ICH_t95 <- round(subset(quantiles_table, reach == "mainstem, MCN to ICH or PRA")$t95, 0)

# first, drop the NAs
ICH_temp_long %>% 
  filter(!is.na(temp)) %>% 
  mutate(date = date(date)) %>% 
  # create a season field
  mutate(month = month(date)) %>% 
  mutate(season = ifelse(month %in% c(1,2,3), "Winter",
                         ifelse(month %in% c(4,5,6), "Spring",
                                ifelse(month %in% c(7,8,9), "Summer",
                                       ifelse(month %in% c(10,11,12), "Fall",
                                              "ERROR"))))) %>% 
  mutate(t50 = NA, t80 = NA, t95 = NA) %>% 
  # let's start in 2005, since that's when our data starts
  subset(year >= 2005) -> ICH_temp_long

# loop to get all windows
for (i in 1:nrow(subset(ICH_temp_long, year <= 2022))){
  ICH_temp_long$t50[i] <- mean(subset(ICH_temp_long, date >= ICH_temp_long$date[i] & date <= ICH_temp_long$date[i]+ days(x = ICH_t50))$temp)
  ICH_temp_long$t80[i] <- mean(subset(ICH_temp_long, date >= ICH_temp_long$date[i] & date <= ICH_temp_long$date[i]+ days(x = ICH_t80))$temp)
  ICH_temp_long$t95[i] <- mean(subset(ICH_temp_long, date >= ICH_temp_long$date[i] & ICH_temp_long$date[i]+ days(x = ICH_t95))$temp)
}

# plot the correlation
ggplot(ICH_temp_long, aes(x = t50, y = t80, color = season)) +
  geom_point() +
  scale_color_tableau(palette = "Tableau 10")

ggplot(ICH_temp_long, aes(x = t50, y = t95, color = season)) +
  geom_point() +
  scale_color_tableau(palette = "Tableau 10")
```
